---
title: "Project_STAT632"
author: "Group 11, Section 2"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(pacman)
p_load(tidyverse, MASS, car, ggplot2,ggcorrplot,stats,corrplot)
```


```{r}
graddata <- read.csv("Admission_Predict.csv")
glimpse(graddata)
```

#Data pre-processing
```{r}
head(graddata)
dim(graddata)

#checking for NA values

gg_miss_var(graddata)

vis_miss(graddata)
```

```{r}
skimr::skim(graddata)
```

```{r}
graddata <- graddata %>%
  rename(gre = GRE.Score,
         srno = Serial.No.,
         toefl = TOEFL.Score,
         unirank = University.Rating,
         chance = Chance.of.Admit)
graddata$unirank <- as.factor(graddata$unirank)
graddata$Research <- as.factor(graddata$Research)

# Excluding the "Serial No." column from the dataset
graddata <- graddata[, -1]
```

### #Summary statistics and EDA:

```{r}
attach(graddata)
summary(graddata)

# Scatterplot matrix
pairs(chance ~ ., data = graddata)
```


```{r}
#Histogram (Continuous Variables)
par(mfrow = c(2, 2))
hist(graddata$chance,main = "Histogram of Chance of Admission",xlab= "Chance of Admission",col = "blue")
hist(graddata$gre,main = "Histogram of GRE",xlab= "GRE score",col = "seagreen")
hist(graddata$toefl,main = "Histogram of TOEFL",xlab= "TOEFL score",col = "violet")
hist(graddata$SOP,main = "Histogram of SOP",xlab= "SOP",col = "lightblue")
hist(graddata$LOR,main = "Histogram of LOR",xlab= "LOR",col = "lightgreen")
hist(graddata$CGPA,main = "Histogram of CGPA",xlab= "CGPA",col = "lightcoral")

#Barplot (Categorical Variables)
par(mfrow = c(2, 2))
ggplot(graddata, aes(x = unirank)) +
  geom_bar(fill = "turquoise",color="blue") +
  labs(x = "University Rating", y = "Count")

ggplot(graddata, aes(x = Research)) +
  geom_bar(fill = "pink",color="blue") +
  labs(x = "Research Experience", y = "Count")
```

```{r}
# Boxplots for GRE score and TOEFL score across different levels of University Rating
ggplot(graddata, aes(x = factor(unirank), y = gre)) +
  geom_boxplot(fill = "skyblue", color = "blue") +
  labs(x = "University Rating", y = "GRE Score") +
  ggtitle("Boxplot of GRE Score by University Rating")

ggplot(graddata, aes(x = factor(unirank), y = toefl)) +
  geom_boxplot(fill = "lightgreen", color = "darkgreen") +
  labs(x = "University Rating", y = "TOEFL Score") +
  ggtitle("Boxplot of TOEFL Score by University Rating")


```

**Relationship b/w Chance of Admission and other variables**

```{r}
# Scatterplot of Chance of Admission by 
boxplot(chance ~ unirank, data = graddata, main = "Chance of Admission by University Rank", xlab = "University Rank", ylab = "Chance of Admission")

boxplot(chance ~ Research, data = graddata, main = "Chance of Admission by Research Experience", xlab = "Research Experience", ylab = "Chance of Admission")

```



**Heatmap showcasing the correlation among the parameters of the data**

```{r}
# Combining Numeric and Factor variables:
combined_data <- cbind(graddata[, sapply(graddata, is.numeric)], 
                        data.frame(sapply(graddata[, sapply(graddata, is.factor)], as.numeric)))

# Calculating correlation matrix
corr_matrix <- cor(combined_data)

# Correlation heatmap:
ggcorrplot(corr_matrix, type = "lower", outline.col = "black", lab = TRUE)
```

**Observations:**

- *As seen in above heatmap 'CGPA' parameter is highly correlated with 'Chance of Admit' whereas 'LOR' parameter is least correlated with other parameters.*

- gre (GRE Scores) and toefl (TOEFL Scores): There is a strong positive correlation of approximately 0.84, indicating that higher GRE scores tend to be associated with higher TOEFL scores.

- gre (GRE Scores) and chance (Chance of Admit): There is a strong positive correlation of approximately 0.80, suggesting that higher GRE scores are associated with a higher chance of admission.

- toefl (TOEFL Scores) and chance (Chance of Admit): Similarly, there is a strong positive correlation of approximately 0.79, indicating that higher TOEFL scores are associated with a higher chance of admission.

- SOP (Statement of Purpose) and LOR (Letter of Recommendation): These variables are positively correlated, with a correlation coefficient of approximately 0.73, indicating that stronger statements of purpose tend to be associated with stronger letters of recommendation.

- CGPA (Undergraduate GPA) and chance (Chance of Admit): There is a strong positive correlation of approximately 0.87, suggesting that a higher undergraduate GPA is associated with a higher chance of admission.





#### Model Building:

```{r}
#Full Model:
model1 <- lm(chance ~ ., data = graddata)
summary(model1)
```
**Interpretation**
- The model explains approximately 80.52% of the variance in the chance of admission.
- Significant predictors of admission include GRE score, TOEFL score, LOR, CGPA, and Research experience.
- Having research experience (Research1) is associated with an increase in the chance of admission.

**full model:**
$\hat{chance} = \hat{\beta_0}+\hat{\beta_1}(gre)+\hat{\beta_2}(toefl)+\hat{\beta_3}(unirank)+\hat{\beta_4}(SOP)+\hat{\beta_5}(LOR)+\hat{\beta_6}(CGPA)+\hat{\beta_6}(Research)$

$$\hat{chance} = −1.2438+0.0017(gre)+0.0031(toefl)−0.0147(unirank2)−0.0093(unirank3)−0.0074(unirank4)+0.0104(unirank5)−0.0026(SOP)+0.0228(LOR)+0.1187(CGPA)+0.0244(Research1)$$



```{r}
par(mfrow = c(1,2))
plot(model1, 1:2)
```


```{r}
shapiro.test(resid(model1))
```

```{r}
#Power Transformation:
boxcox(model1, lambda = seq(1, 3, 0.05))
summary(powerTransform(model1))
```

```{r}
#using transformed response

model2 <- lm(chance^2.61 ~ gre + toefl + unirank + SOP + LOR + CGPA + Research, data = graddata)
summary(model2)

par(mfrow = c(2,2))
plot(model2)
```

```{r}
model3 <- lm(chance^2.61 ~ gre + toefl + unirank + LOR + CGPA + Research, data = graddata)
summary(model3)
AIC(model3)
```

```{r}
shapiro.test(resid(model3))
```

```{r}
plot(model3, 1:2)
```

```{r}
#Example prediction :
newdata <- data.frame(gre = 324, toefl = 116, SOP = 4.5, LOR = 4.5, Research = factor(0), CGPA = 7.8, unirank = factor(4))

predict(model3, newdata)
```

```{r}
model3_step <- step(model3)
summary(model3_step)
```

```{r}
plot(model3_step, 1:2)
```

```{r}
AIC(model1, model2, model3, model3_step)
#final model
```

Hence our final model(model3_step), a linear regression, includes predictors such as GRE score, TOEFL score, University Rating, Statement of Purpose (SOP), Letter of Recommendation (LOR), CGPA, and Research Experience. This model explains approximately 80.52% of the variance in the chance of admission. Significant predictors of admission include GRE score, TOEFL score, LOR, CGPA, and Research experience.



























